~\\ \Large\bf Data Intensive processing with iRODS and the middleware CiGri for the Whisper project

### 2014-08-29 articles/2014/whisper/aricleEbauche/Draft_BriandBzeznikIrodsUserMeeting2014.tex
very small change
add References, cite particule and genomics projet and science article

\begin{document}
\long\def\/*#1*/{}
\date{}%{\today}
\title{\huge{DRAFT} \\ ~\\ \Large\bf Data Intensive processing with iRODS and the middleware CiGri for the Whisper project}
\author{\begin{tabular}[t]{c@{\extracolsep{8em}}c}
	Briand Xavier\footnote{Cnrs, Isterre, Whisper, email xav.briand@gmail.com}  & Bzeznik Bruno\footnote{Ciment, Universit\'e Joseph Fourier, Technical header of Ciment} \\
\end{tabular}}
\maketitle
%\thispagestyle{empty}
\subsection*{Abstract}
{\em
Like many projects in science of the universe, the seismological project Whisper is faced with massive data processing.
This leads to specific IT software for the project as well as suitable IT infrastructure. We present here both aspects.

@@

This is the partnership between these two IT components that has enabled a data-intensive processing and, also,  permits to the Whisper project to bring new scientific results.
}
~\\ \\
{\bf Keywords}:  Data-Intensive, grid computing, distributed storage, Seismic Noise, Whisper, Cigri, Irods.

@@

The University of Grenoble offers, thanks to the High Performance Computing (HPC) centre \emph{Ciment}, this kind of environement with the distributed file system \emph{Irods}
and the middelware \emph{CiGri}.% (Ciment platform).

@@

\begin{center}%[h]
\centering
\captionof{figure}{\label{schemaNodeStorageCiment} A storage node}
\includegraphics[width=5cm]{schemaNodeStorageCiment.png}
\end{center}
iRODS nodes are running Debian GNU/Linux with Kanif \footnote{see http://taktuk.gforge.inria.fr/kanif/} for easy synchronisation of the system administration.
CIMENT has set up a web interface where the user can easily check the status of the ressources (figure \ref{schemaCimentStorageStatus}).

@@

Whisper is one of the projects that have made it possible to ensure that the seismic noise brought new observable.
This permits to carry out several scientific results including imaging and monitoring. 
Concerning monitoring, further studies provide news results about sligth variations of seismics waves induce by earthquakes.
Many articles are in part due to this project as well as
several post-docs and PhD (see \cite{key:CFLC2010, key:BPCPBR, key:HCM2014, key:OPLC2013, key:PCP, key:MPPVCS}) and also whisper.obs.ujf-grenoble.fr, rubric publication).

@@

We illustrate with the figure \ref{JapanVelocity}, one of the analysis of the correlations that represents an image
of change of the velocity of seismic waves.
\begin{center}%[h]
\centering
\captionof{figure}{\label{JapanVelocity}}
\includegraphics[width=8cm]{JapanVelocity.png}
\end{center}

@@

With the Whisper project, as with some few other projects, we had the opporunity to test and improve the CIMENT IT infrastructure for a data-intensive case. In this kind of projects, data management and input/output data flows are a big part of the process regarding the actual computing time for the analysis. With a grid of several thousands cpu-cores, such a project may act as a real distributed attack against our storage infrastructures! So, as a first consequence of the deployment of such computing jobs, we had to implement new "limiting" functionnalities into our infrastructure softwares or configurations. For example, CiGri is able to limit the number of submitted jobs for a given campaign on a given cluster if we know that the jobs concurrency can overload our Irods infrastructure. As another example, iRODS may be configured to limit the number of simultaneous connections. But in this case, the i-command returns with an error and the job may be aborted. There's a "--retry" optionnal argument, but with a high level of concurrency, it increases the load and may completely exhaust the servers with a ton of retry queries. We then developped a wrapper for the i-commands, called secure-i-commands that implements a retry on some specific error cases  with an incremental delay to prevent from flooding the irods servers. Of course, this can be improved as this leads to a waste of computing resources because they are reserved and not used while waiting. But this is at least a security improvement in case the highler level decision processes does not work as expected because of their high complexity.

@@

Lot of improvements that involve automatization of the processing, and also concerning the scientific workflow have constantly been made.

We may say that Whisper+CiGri+iRODS is a great success!

@@

\bibitem{key:BPCPBR}
P Bou\'e, P Poli, M Campillo, H Pedersen, X Briand, P Roux,
Teleseismic correlations of ambient seismic noise for deep global imaging of the Earth.
{\em Geophysical Journal International 194 (2), 844-848.}

@@

\bibitem{key:CHR2009}
Beno\^it Claudel, Guillaume Huard, Olivier Richard,
TakTuk, Adaptive Deployment of Remote Executions.
{\em Proceedings of the International Symposium on High Performance Distributed Computing (HPDC) (2009) 91-100.}

@@

@@

>>>>

\begin{document}
\long\def\/*#1*/{}
\date{}%{\today}
%\title{\huge{DRAFT} \\ ~\\ \Large\bf Data Intensive processing with iRODS and the middleware CiGri for the Whisper project}
\title{ ~\\ \Large\bf Data Intensive processing with iRODS and the middleware CiGri for the Whisper project}

\author{\begin{tabular}[t]{c@{\extracolsep{8em}}c}
	Briand Xavier\footnote{Cnrs, Isterre, Whisper, email xav.briand@gmail.com}  & Bzeznik Bruno\footnote{Ciment, Universit\'e Joseph Fourier, Technical header of Ciment} \\
\end{tabular}}
\maketitle
%\thispagestyle{empty}
\subsection*{Abstract}
{\em
Like many projects in science of the universe, the seismological project Whisper is faced with massive data processing.
This leads to specific IT software for the project as well as suitable IT infrastructure. We present here both aspects.

@@

This is the partnership between these two IT components that has enabled a data-intensive processing and, also,  permits to the Whisper project to bring new scientific results.
}
~\\ \\
{\bf Keywords}:  Data-Intensive, Grid computing, Distributed storage, Seismic Noise, Whisper, Cigri, Irods.

@@

The University of Grenoble offers, thanks to the High Performance Computing (HPC) centre \emph{Ciment}, this kind of environement with the data management system \emph{Irods}
and the middelware \emph{CiGri}.% (Ciment platform).

@@

\begin{center}%[h]
\centering
\captionof{figure}{\label{schemaNodeStorageCiment} A storage node}
\includegraphics[width=5cm]{schemaNodeStorageCiment.png}
\end{center}
iRODS nodes are running Debian GNU/Linux with Kanif \footnote{see http://taktuk.gforge.inria.fr/kanif/} \cite{key:CHR2009} for easy synchronisation of the system administration.
CIMENT has set up a web interface where the user can easily check the status of the ressources (figure \ref{schemaCimentStorageStatus}).

@@

Whisper is one of the projects that have made it possible to ensure that the seismic noise brought new observable.
This permits to carry out several scientific results including imaging and monitoring. 
Concerning monitoring, further studies provide news results about sligth variations of seismics waves induce by earthquakes.
Many articles are in part due to this project as well as
several post-docs and PhD (see \cite{key:CFLC2010, key:BPCPBR, key:HCM2014, key:OPLC2013, key:PCP, key:MPPVCS, key:RCRZCSKCCWD, key:RBCSP, key:BCTASBEM}) and also whisper.obs.ujf-grenoble.fr, rubric publication).

@@

We illustrate with the figure \ref{JapanVelocity}, one of the analysis of the correlations that represents an image
of change of the velocity of seismic waves in Japan (the giant 2011 Tohoku earthquake, see \cite{key:BCTASBEM}).
\begin{center}%[h]
\centering
\captionof{figure}{\label{JapanVelocity}}
\includegraphics[width=8cm]{JapanVelocity.png}
\end{center}

@@

With the Whisper project, as with some few other projects, we had the opporunity to test and improve the CIMENT IT infrastructure for a data-intensive case. 
Note also that the data-grid environment is used by many other scientifics projects such that research in Particle Physics \cite{key:BB} or genomics research \cite{key:MLBDJ}.
For this kind of project, data management and input/output data flows are a big part of the process regarding the actual computing time for the analysis. With a grid of several thousands cpu-cores, such a project may act as a real distributed attack against our storage infrastructures! So, as a first consequence of the deployment of such computing jobs, we had to implement new "limiting" functionnalities into our infrastructure softwares or configurations. For example, CiGri is able to limit the number of submitted jobs for a given campaign on a given cluster if we know that the jobs concurrency can overload our Irods infrastructure. As another example, iRODS may be configured to limit the number of simultaneous connections. But in this case, the i-command returns with an error and the job may be aborted. There's a "--retry" optionnal argument, but with a high level of concurrency, it increases the load and may completely exhaust the servers with a ton of retry queries. We then developped a wrapper for the i-commands, called secure-i-commands that implements a retry on some specific error cases  with an incremental delay to prevent from flooding the irods servers. Of course, this can be improved as this leads to a waste of computing resources because they are reserved and not used while waiting. But this is at least a security improvement in case the highler level decision processes does not work as expected because of their high complexity.

@@

Lot of improvements that involve automatization of the processing, and also concerning the scientific workflow have constantly been made.
We may say that Whisper+CiGri+iRODS is a great success!

\section{Acknowledgements}
We thanks Emmanuel Chaljub, the OAR team and the Ciment team as well as the Whisper project and the Verce project for all their  very useful support.

@@

\bibitem{key:BB}
C. Biscarat, B. Bzeznik.
Synergy between the CIMENT tier-2 HPC centre and the HEP community at LPSC in Grenoble (France)
{\em J. Phys.: Conf. Ser. 513 032008, 2014}

\bibitem{key:BPCPBR}
P Bou\'e, P Poli, M Campillo, H Pedersen, X Briand, P Roux.
Teleseismic correlations of ambient seismic noise for deep global imaging of the Earth.
{\em Geophysical Journal International 194 (2), 844-848.}

\bibitem{key:BCTASBEM}
Brenguier, F., M. Campillo, T. Takeda, Y. Aoki, N. M. Shapiro, X. Briand, K. Emoto, H. Miyake (2014).
Mapping pressurized volcanic fluids from induced crustal seismic velocity drops.
{\em Science, 345, 80-82, 2014.}

@@

\bibitem{key:CHR2009}
Beno\^it Claudel, Guillaume Huard, Olivier Richard.
TakTuk, Adaptive Deployment of Remote Executions.
{\em Proceedings of the International Symposium on High Performance Distributed Computing (HPDC) (2009) 91-100.}

@@

\bibitem{key:MLBDJ}
M\"unkem\"uller T., Lavergne S., Bzeznik B., Dray S., Jombart T.,
Schiffers K., Thuiller W. 
How to measure and test phylogenetic signal. Methods in Ecology and Evolution, 3: 743-756, 2012.

@@

\bibitem{key:RCRZCSKCCWD}
D. Rivet, M. Campillo, M. Radiguet, D. Zigone, V. Cruz-Atienza, N. M. Shapiro, V. Kostoglodov, N. Cotte,
G. Cougoulat, A. Walpersdorf and E. Daub
Seismic velocity changes, strain rate and non-volcanic tremors during the 2009-2010 slow slip event in Guerrero, Mexico. 
{\em Geophys. J. Int. 2014;196:447-460. }

\bibitem{key:RBCSP}
Rivet D., Brenguier F., Clarke D., Shapiro N.M., Peltier A. 
Long-term dynamics of Piton de la Fournaise volcano from 13 years of seismic velocity changes measurements and GPS observation.
{\em J. of Geophys. Res. 2014; submitted.}
